Language-Based Security for Low-Level MPC

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

For full details of the GMW protocol the reader is referred to
\cite{evans2018pragmatic}. Our implementation libary is shown in
Figure \ref{fig-gmw}, and includes encoding functions, where
input secrets are split into shares, $\eand$ and $\exor$ gate
functions, and a decoding function. Note that $\exor$ requires
no interaction between parties, while conjunction necessitates
1-of-4 oblivious transfer. The gate computation is
done entirely in secrect, and the decoding function
is where the declassification occurs-- both parties reveal
their shares of the final gate output $\macgv{z}$.

@@

As discussed above and in related work \cite{8429300}, probabilistic
separation conditional on certain variables-- e.g., secret inputs or
public outputs-- is a key mechanism for reasoning about MPC protocol
security. Following \cite{barthe2019probabilistic}, we define a
conditional separation relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to
mean that $X_2$ and $X_3$ are indpendent in $\pmf$ conditionally on
any assignment of values to $X_1$-- i.e., conditionally on any $\store
\in \mems(X_1)$. Another key concept needed especially for reasoning
about circuits is conditional determinism. For example, if $\macgv{z}$
is an output of an internal gate, it will definitely be computed using
random variables, however, it \emph{should} be determistic under any
set of input secrets $S$, since we assume that $\idealf$ is
deterministic. Conditional uniformity is also important, since the
gradual release property of many protocols means that messages appear
in a uniform distribution to the adversary.
\begin{definition}[Conditioning Properties] 
  Give $\pmf$ and $X_1,X_1,X_3 \subseteq \dom{pmf}$, 
  we write:
  \begin{itemize}
  \item $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$ we have
    $\pmf(\store|\store') = \pmf(\store_{X_2}|\store') *  \pmf(\store_{X_3}|\store')$.
  \item $\conddetx{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ there exists 
    $\store \in \mems(X_2)$ such that $\pmf(\store|\store') = 1$.
  \item $\condunix{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ and
    $\store \in \mems(X_2)$ we have
    $\pmf(\store|\store') = 1/p^{|X_2|}$.
  \end{itemize}
\end{definition}

Given these definitions, we can formulate an invariant for circuit
computation with respect to internal gates as follows. It says that
the output of any gate is deterministic given inputs $S$, and
conditionally on $S$ corrupt views are always in a uniform random
disrtibution (pure noise), while output $\macgv{\mesg{z}}$ remains
separable from corrupt views and both shares of
$\macgv{\mesg{z}}$. This last condition (incidentally missing from PSL
due to the reliance on more recent innovations in conditioning logic
\cite{li2023lilac}) is critical since those shares will in fact be
revealed if $\macgv{\mesg{z}}$ is decoded as the circuit output.
\begin{lemma}[GMW Invariant]
  \label{lemma-gmwinvariant}
  Given:
  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
  Then all of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
  \begin{enumerate}
  \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
  \item $\condunix{\progtt(\prog)}{S}{M_C}$
  \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
  \end{enumerate}
\end{lemma}
To prove this, we can formulate and automatically prove local,
gate-level versions of the invariant. This serves as a proof tactic
that simplifies the proof of the GMW invariant. 
\begin{lemma}[And Gate Tactic]
  \label{lemma-gmwtactic}
  %Define:
  %$$
  %\begin{array}{c}
  %  \prog_{i} \defeq \\
  %  \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
  %  \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
  %\end{array}
  %$$
  Given:
  $$
  \begin{array}{c}
  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
  (\prog,\mesg{z})
  \end{array}
  $$
  Then all of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
  \begin{enumerate}
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
\end{lemma}
\begin{proof}
Verified automatically using techniques described in Section \ref{section-bruteforce}.  
\end{proof}

>>>>

For full details of the GMW protocol the reader is referred to
\cite{evans2018pragmatic}. Our implementation library is shown in
Figure \ref{fig-gmw}, and includes encoding functions, where
input secrets are split into shares, $\eand$ and $\exor$ gate
functions, and a decoding function. Note that $\exor$ requires
no interaction between parties, while conjunction necessitates
1-of-4 oblivious transfer. The gate computation is
done entirely in secret, and the decoding function
is where the declassification occurs-- both parties reveal
their shares of the final gate output $\macgv{z}$.

@@

As discussed above and in related work \cite{8429300}, probabilistic
separation conditional on certain variables-- e.g., secret inputs or
public outputs-- is a key mechanism for reasoning about MPC protocol
security. Following \cite{barthe2019probabilistic}, we define a
conditional separation relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to
mean that $X_2$ and $X_3$ are independent in $\pmf$ conditionally on
any assignment of values to $X_1$-- i.e., conditionally on any $\store
\in \mems(X_1)$. Another key concept needed especially for reasoning
about circuits is conditional determinism. For example, if $\macgv{z}$
is an output of an internal gate, it will definitely be computed using
random variables, however, it \emph{should} be deterministic under any
set of input secrets $S$, since we assume that $\idealf$ is
deterministic. Conditional uniformity is also important, since the
gradual release property of many protocols means that messages appear
in a uniform distribution to the adversary.
\begin{definition}[Conditioning Properties] 
  Give $\pmf$ and $X_1,X_1,X_3 \subseteq \dom{pmf}$, 
  we write:
  \begin{itemize}
  \item $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$ we have
    $\pmf(\store|\store') = \pmf(\store_{X_2}|\store') *  \pmf(\store_{X_3}|\store')$.
  \item $\conddetx{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ there exists 
    $\store \in \mems(X_2)$ such that $\pmf(\store|\store') = 1$.
  \item $\condunix{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ and
    $\store \in \mems(X_2)$ we have
    $\pmf(\store|\store') = 1/p^{|X_2|}$.
  \end{itemize}
\end{definition}

Given these definitions, we can formulate an invariant for circuit
computation with respect to internal gates as follows. It says that
the output of any gate is deterministic given inputs $S$, and
conditionally on $S$ corrupt views are always in a uniform random
distribution (pure noise), while output $\macgv{\mesg{z}}$ remains
separable from corrupt views and both shares of
$\macgv{\mesg{z}}$. This last condition (incidentally missing from PSL
due to the reliance on more recent innovations in conditioning logic
\cite{li2023lilac}) is critical since those shares will in fact be
revealed if $\macgv{\mesg{z}}$ is decoded as the circuit output.
\begin{lemma}[GMW Invariant]
  \label{lemma-gmwinvariant}
  Given:
  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
  Then all of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
  \begin{enumerate}
  \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
  \item $\condunix{\progtt(\prog)}{S}{M_C}$
  \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
  \end{enumerate}
\end{lemma}
To prove this, we can formulate and automatically prove local,
gate-level versions of the invariant. This serves as a proof tactic
that simplifies the proof of the GMW invariant. 
\begin{lemma}[And Gate Tactic]
  \label{lemma-gmwtactic}
  %Define:
  %$$
  %\begin{array}{c}
  %  \prog_{i} \defeq \\
  %  \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
  %  \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
  %\end{array}
  %$$
  Given:
  $$
  \begin{array}{c}
  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
  (\prog,\mesg{z})
  \end{array}
  $$
  Then all of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
  \begin{enumerate}
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
\end{lemma}
\begin{proof}
Verified automatically using techniques described in Section \ref{section-bruteforce}.  
\end{proof}

### 2024-05-22 ppdp24/intro.tex
checkpoint

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{evans2018pragmatic}.  Related proof methods are well-studied
\cite{Lindell2017} but mostly manual. Somewhat independently, a
significant body of work in programming languages has focused on
definition and enforcement of confidentiality and integrity
\emph{hyperproperties} \cite{10.5555/1891823.1891830} such as
noninterference and gradual release
\cite{4223226,sabelfeld2009declassification}. Following a tradition of
connecting cryptographic and PL-based security models
\cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
recognized connections between MPC security models and hyperpoperties
of, e.g., noninterference \cite{8429300}, and even leveraged these
connections to enforce MPC security through mechanisms such as
security types \cite{10.1145/3453483.3454074}. Major benefits of this
connection in an MPC setting include better language abstractions for
defining protocols and for mechanisation and even automation of
security proofs.  The goal of this paper is to develop a PL model for
defining a variety of low-level probabilistic MPC protocols, to
formulate a collection of confidentiality and integrity
hyperproperties for our model model with familiar information flow
analogs, and to show how these properties can be leveraged for
improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria
\cite{rastogi2014wysteria} and Viaduct \cite{10.1145/3453483.3454074}
are designed to provide effective programming of full
applications. These language designs incorporate sophisticated
verified compilation techniques such as orchestration
\cite{viaduct-UC} to guarantee high-level security properties, and
they rely on \emph{libraries} of low-level MPC protocols, such as
binary and arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encrypytion,
and must be verified by hand. So, low-level MPC
programming and protocol verification remains a distinct challenge and both
critical to the general challenge of PL design for MPC and
complementary to high-level language design.

@@

As mentioned above, several high-level languages have been developed
for writing MPC applications, and frequently exploit the connection
between hyperproperties and MPC security. Previous work on analysis
for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. These two lines of work were focused on passive
security. The Viaduct language
\cite{10.1145/3453483.3454074} has a well-developed
information flow type system that automatically enforces both
confidentiality and integrity through hyperproperties such as robust
declassification, in addition to rigorous compilation guarantees
through orchestration \cite{viaduct-UC}. However, these high level
languages lack probabilistic features and other abstractions of
low-level protocols, the implementation and security of which are
typically assumed as a selection of library compnonents.

Various related low-level languages with probabilistic features have
also been developed. The $\lambda_{\mathrm{obliv}}$ language
\cite{darais2019language} uses a type system is used to automatically
enforce so-called probabilistic trace obliviousness.  But similar to
previous work on oblivious data structures \cite{10.1145/3498713},
obliviousness is related to pure noninterference, not the relaxed form
related to passive MPC security. The Haskell-based security type
system in \cite{6266151} enforces a version of noninterence that is
sound for passive security, but does not verify the correctness of
declassifications and does not consider malicious security. And
properties of real/ideal passive and malicious security for a
probabilistic language have been formulated in EasyCrypt
\cite{8429300}-- though their proof methods, while mechanized, are
fully manual, and their formulation of malicious security is not as
clearly related to robust declassification as is the one we present in
Section \ref{section-hyper}.

>>>>

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{evans2018pragmatic}.  Related proof methods are well-studied
\cite{Lindell2017} but mostly manual. Somewhat independently, a
significant body of work in programming languages has focused on
definition and enforcement of confidentiality and integrity
\emph{hyperproperties} \cite{10.5555/1891823.1891830} such as
noninterference and gradual release
\cite{4223226,sabelfeld2009declassification}. Following a tradition of
connecting cryptographic and PL-based security models
\cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
recognized connections between MPC security models and hyperpoperties
of, e.g., noninterference \cite{8429300}, and even leveraged these
connections to enforce MPC security through mechanisms such as
security types \cite{10.1145/3453483.3454074}. Major benefits of this
connection in an MPC setting include better language abstractions for
defining protocols and for mechanization and even automation of
security proofs.  The goal of this paper is to develop a PL model for
defining a variety of low-level probabilistic MPC protocols, to
formulate a collection of confidentiality and integrity
hyperproperties for our model model with familiar information flow
analogs, and to show how these properties can be leveraged for
improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria
\cite{rastogi2014wysteria} and Viaduct \cite{10.1145/3453483.3454074}
are designed to provide effective programming of full
applications. These language designs incorporate sophisticated
verified compilation techniques such as orchestration
\cite{viaduct-UC} to guarantee high-level security properties, and
they rely on \emph{libraries} of low-level MPC protocols, such as
binary and arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encryption,
and must be verified by hand. So, low-level MPC
programming and protocol verification remains a distinct challenge and both
critical to the general challenge of PL design for MPC and
complementary to high-level language design.

@@

As mentioned above, several high-level languages have been developed
for writing MPC applications, and frequently exploit the connection
between hyperproperties and MPC security. Previous work on analysis
for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. These two lines of work were focused on passive
security. The Viaduct language
\cite{10.1145/3453483.3454074} has a well-developed
information flow type system that automatically enforces both
confidentiality and integrity through hyperproperties such as robust
declassification, in addition to rigorous compilation guarantees
through orchestration \cite{viaduct-UC}. However, these high level
languages lack probabilistic features and other abstractions of
low-level protocols, the implementation and security of which are
typically assumed as a selection of library components.

Various related low-level languages with probabilistic features have
also been developed. The $\lambda_{\mathrm{obliv}}$ language
\cite{darais2019language} uses a type system is used to automatically
enforce so-called probabilistic trace obliviousness.  But similar to
previous work on oblivious data structures \cite{10.1145/3498713},
obliviousness is related to pure noninterference, not the relaxed form
related to passive MPC security. The Haskell-based security type
system in \cite{6266151} enforces a version of noninterference that is
sound for passive security, but does not verify the correctness of
declassifications and does not consider malicious security. And
properties of real/ideal passive and malicious security for a
probabilistic language have been formulated in EasyCrypt
\cite{8429300}-- though their proof methods, while mechanized, are
fully manual, and their formulation of malicious security is not as
clearly related to robust declassification as is the one we present in
Section \ref{section-hyper}.
